---
title: "p8105_hw2_yl5214"
output: github_document
---

```{r}
library(tidyverse)
```

## Question 2
## Read and clean the Mr. Trash Wheel sheet
```{r}
Mr_trash_df =
  readxl::read_excel('data/202309 Trash Wheel Collection Data.xlsx', sheet = 1, range = "A2:N586" ) |> 
  mutate( trash_wheel= 'Mr_trash_wheel') |> 
 janitor::clean_names() |> 
  mutate(
    homes_powered = weight_tons*500/30
    ) 
Mr_trash_df$year = as.character(Mr_trash_df$year)
```


## Import, clean, and organize the data for Professor Trash Wheel and Gwynnda
```{r}
professor_trash_df =
  readxl::read_excel('data/202309 Trash Wheel Collection Data.xlsx', sheet = 2, range = "A2:M108" ) |> 
  mutate( trash_wheel= 'professor_trash_wheel') |> 
 janitor::clean_names() |> 
  mutate(
    homes_powered = weight_tons*500/30
  )
professor_trash_df$year = as.character(professor_trash_df$year)

gwynnda_df =
  readxl::read_excel('data/202309 Trash Wheel Collection Data.xlsx', sheet = 4, range = "A2:L157" ) |>
  mutate( trash_wheel= 'gwynnda_trash_wheel') |>
 janitor::clean_names() |> 
  mutate(
    homes_powered = weight_tons*500/30
  ) 
gwynnda_df$year = as.character(gwynnda_df$year)

```

## Gwynnda in July of 2021
```{r}
total_cig =filter(gwynnda_df, year =='2021', month =='July')
```


## Combine these with the Mr. Trash Wheel dataset to produce a single tidy dataset
```{r}
combine_df =
  bind_rows(Mr_trash_df, professor_trash_df, gwynnda_df) |> 
  janitor::clean_names() |> 
relocate(trash_wheel)
```

## Description
There are `r nrow(combine_df)` rows and `r ncol(combine_df)` columns in the merged dataset. The key variables are `r names(combine_df)`. The total weight of trash collected by Professor Trash Wheel is `r sum(professor_trash_df$ weight_tons)` The total number of cigarette butts collected by Gwynnda in July of 2021 is `r sum(total_cig$cigarette_butts)`








