---
title: "p8105_hw2_yl5214"
output: github_document
---

```{r}
library(tidyverse)
```

## Question 2
## Read and clean the Mr. Trash Wheel sheet
```{r}
Mr_trash_df =
  readxl::read_excel('data/202309 Trash Wheel Collection Data.xlsx', sheet = 1, range = "A2:N586") |> 
  mutate( trash_wheel= 'Mr_trash_wheel') |> 
 janitor::clean_names() |> 
  drop_na (dumpster) |> 
  mutate(
    homes_powered = weight_tons*500/30
    ) 
Mr_trash_df$year = as.character(Mr_trash_df$year)
```


## Import, clean, and organize the data for Professor Trash Wheel and Gwynnda
```{r}
professor_trash_df =
  readxl::read_excel('data/202309 Trash Wheel Collection Data.xlsx', sheet = 2, range = "A2:M108") |> 
  mutate( trash_wheel= 'professor_trash_wheel') |> 
 janitor::clean_names() |>
  drop_na (dumpster) |> 
  mutate(
    homes_powered = weight_tons*500/30
  )
professor_trash_df$year = as.character(professor_trash_df$year)

gwynnda_df =
  readxl::read_excel('data/202309 Trash Wheel Collection Data.xlsx', sheet = 4, range = "A2:L157", na ="") |>
  mutate( trash_wheel= 'gwynnda_trash_wheel') |>
 janitor::clean_names() |> 
  drop_na (dumpster) |> 
  mutate(
    homes_powered = weight_tons*500/30
  ) 
gwynnda_df$year = as.character(gwynnda_df$year)

```

## Gwynnda in July of 2021
```{r}
total_cig =filter(gwynnda_df, year =='2021', month =='July')
```


## Combine these with the Mr. Trash Wheel dataset to produce a single tidy dataset
```{r}
combine_df =
  bind_rows(Mr_trash_df, professor_trash_df, gwynnda_df) |> 
  janitor::clean_names() |> 
relocate(trash_wheel)
```

## Description
There are `r nrow(combine_df)` rows and `r ncol(combine_df)` columns in the merged dataset. The key variables are `r names(combine_df)`. The total weight of trash collected by Professor Trash Wheel is `r sum(professor_trash_df$ weight_tons)` The total number of cigarette butts collected by Gwynnda in July of 2021 is `r sum(total_cig$cigarette_butts)`

## Question 3


## Import, clean, and tidy the dataset of baseline demographics
```{r}
mci_baseline=
  read_csv("data/MCI_baseline.csv", 
           skip= 1)|> 
  janitor:: clean_names() |> 
  filter ( current_age < age_at_onset| age_at_onset == "." ) |> 
    mutate(
       sex = case_match(
         sex,
        0 ~ "female",
        1 ~ "male"
        )
) |> 
       mutate(
        apoe4 =case_match(
          apoe4,
        0 ~ " APOE4 non-carrier",
        1 ~ "APOE4 carrier" 
        ) 
) 


```

##Discuss important steps in the import process and relevant features of the dataset.
First, I delete a row at the top of the dataset because the first row does not contain the variable names for each column. Then, I remove the participants who do not meet the stated inclusion criteria by filtering current age less than the age at onset or the age at onset is ".". Then, I change the sex status from 1, 0 to male, female by using the mutate statement. Also, I change the apoe4 carrier status from 1, 0 to APOE4 carrier, APOE4 non-carrier by using mutate statement.

##How many participants were recruited, and of these how many develop MCI?
Those who develop MCI
```{r}
baseline2= filter(mci_baseline, age_at_onset != '.')
```
`r nrow(mci_baseline)` participants were recruited, and of these `r nrow(baseline2)` develop MCI.
##What is the average baseline age? 
The average baseline age is `r mean(mci_baseline$current_age)` 
## What proportion of women in the study are APOE4 carriers?
```{r}
proportion=
  sum(mci_baseline$sex == "female" & mci_baseline$ apoe4 == "APOE4 carrier") / sum(mci_baseline$sex == "female")
```
`r proportion` women in the study are APOE4 carriers.

##Import, clean, and tidy the dataset of longitudinally observed biomarker values
```{r}
mci_amyloid=
  read_csv("data/mci_amyloid.csv", 
           skip= 1)|> 
  janitor:: clean_names() |> 
  pivot_longer(
   baseline: time_8,
   names_to ='time',
   values_to = 'number',
   names_prefix ='time_'
   )|>
     mutate(
       time = replace(time, time== 'baseline', '0')
     )

```

##Comment on the steps on the import process and the features of the dataset





